Efficient and Robust Plane-Based Boolean Operations on N Meshes

### 2016-12-25 paper.tex
before chagne

\IEEEtitleabstractindextext{
\begin{abstract}
  Constructive solid geometry (CSG) is widely used for computer aided design and manufacturing. However, boolean operations, which compute the boundary representation of CSG, have been suffered from robustness and efficiency problem for more than three decades. We propose a fast and exact boolean operation method which is unconditionally robust for regular set solids. We notice that among previous works,boolean methods using plane representations (P-reps) of meshes are easy to achieve robustness, but hard to be efficient. On the other hand, methods based on boundary representation (B-reps) of meshes are fast but not robust unless exact arithmetic is applied. Therefore, we choose to combine P-reps with B-reps, which allows us to take the advantages of both. Our method is globally based on B-reps for its efficiency, and P-rep geometry is applied locally to ensure robustness. Comparison experiments with state-of-the-art techniques show that our method is unconditionally exact and robust, while is competitively efficient as non-robust methods.
\end{abstract}

@@

@@

Another thing needs to be noticed is that there will be repetitive detection of edge intersections. For example, in Fig. \ref{fig:twin}a, the same intersection on $t_1$ will be detected twice because $t_1$ intersect both $t_2$ and $t_2^{\star}$ in exactly the same intersection. We solve this duplication together with other interactions among different intersections in \S\ref{sec:tessellation}. We also call such $t_2$ and $t_2^{\star}$ as \emph{companion triangles} in an edge intersection because they share the same edge intersection. This concept will be referred again in the discussion of coplanar cases.

@@

\begin{figure}[t]
\centering
\includegraphics[width=3.5in]{boolean-03}
\caption{a) Coplanar cases, $t_a$ and $t_b$ intersect in 2D, dividing each other into two areas---a convex overlapping area and exclusive parts. b) The possible configurations of the companion faces. The blue triangles are from the same mesh. For $\mathcal{I}_1$, the companion triangle $t_1$ is coplanar with the $t-a$. Therefore, the overlap areas do not need to be split by the yellow edge during tessellation, thus $\mathcal{I}_1$ is invalid. For $\mathcal{I}_2$ and $\mathcal{I}_3$, the companion faces are not coplanar, so these intersection can be recorded during intersection test between $t_a$ and $t_1$ and $t_2$. In any configurations, their is no need to record the intersection between $t_a$ and $t_b$.}
\label{fig:coplanar}
\end{figure}

Consider two triangle faces $t_1$ and $t_2$ intersect within a common plane. Both $t_1$ and $t_2$ will divide each other into two areas---a convex overlapping area and exclusive parts (Fig. \ref{fig:coplanar}a). Apparantly, if we tessellate both $t_1$ and $t_2$ according to the boundary of overlapping area, we can guarantee that the tessellated meshes are intersection-free. Many previous methods \cite{feito2013fast,zhou2016mesh} use this method to guarantee topology correctness. However, in our method, we do not test whether $t_1$ and $t_2$ really intersect in 2D once we find they are coplanar. We treat coplanar situations as they do not intersect at all. In this way, we simplify our method, making it more robust and fast, while doing no harm to the topology correctness.

@@

However, if the companion triangles are also coplanar, neither of the two triangles will record the intersection. Fortunately, in this case, the intersection is not valid. The word `valid` means the intersections will appear as an edge in the final mesh, thus is necessary to put into tessellation. Supposing the intersection $\mathcal{I}$  is on triangle $t_1$, this definition equals to:
\begin{equation}
% = f(\bm{\Lambda}(\bm{s}_i))
\lambda_f(\bm{x}_i) \neq constant, \bm{x}_i \in \bm{U}(\mathcal{I}) \cup t_1
\end{equation}
, where $f$ is the CSG function, and $\bm{U}(\mathcal{I})$ is the the neighborhood of $\mathcal{I}$. This inequation indicates the two sides of $\mathcal{I}$ have different classification and must be split according to $\mathcal{I}$ during tessellation. However, in most time, the validation of intersection is not known until classification stage, because we have to know the indicator of each point in $\bm{U}(\mathcal{I}) \cup t_1$. Fortunately, if companion faces are both coplanar with $t_1$, both sides of $\mathcal{I}$ have the same indicator $on$ and thus the same classification. It means $\mathcal{I}$ is invalid and we do not need to record it.

@@

\label{sec:tessellation}
After intersection computation, tessellation is performed on each intersecting triangle face and generate our intersection-free meshes. We call this stage as deferred tessellation because the tessellation happens until all valid intersections are computed.

Ogayar-Anguita et. al. \cite{ogayar2015deferred} use Constrained Delaunay Triangulation (CDT) to perform deferred tessellation. They treat triangle faces as convex triangulation zones and intersections as constraints. Our first try is to implement a robust plane-based CDT. However, CDT algorithms \cite{chew1989constrained,preparata2012computational} require coordinates projection or intersection detection between arbitrary connections of vertices within triangulation zone, which are difficult to be implemented under P-reps. From another perspective, CDT contains geometry constructions---new edges are generated to guarantee each subface is a triangle, which does not fit P-reps. Moreover, when there are more than two primitives, simply applying CDT omits the complexity of intersections between meshes. The intersections may intersect each other, introducing new vertices and splitting original intersection line segments (Fig. \ref{fig:iisect}a). This breaks the assumption of most CDT algorithms and may lead to incorrect output topology.

\begin{figure}[t]
\centering
\includegraphics[width=3.5in]{boolean-04}
\caption{a) Different colors indicate the intersections are from different meshes. The blue intersections and the red ones intersect at a point. Also, the yellow intersections overlap with the red ones. b) After refinement, we introduce a new vertex $a$, and merge overlapping intersections into a new one $\mathcal{I}_b$. c) Our tessellation does not guarantee to obtain triangles. d) If triangulation is performed, new edges (gray ones) are introduced and we cannot give the precise plane-based representation for these new edges.}
\label{fig:iisect}
\end{figure}

To solve these problems, we first perform intersection refinement to eliminate any crossing or overlapping between intersections. After that, we perform conservative tessellation based on \emph{tess-graph}, ensuring unconditional topology correctness. An tess-graph is a graph-like description of intersections on a certain face. The word `conservative` means we . This is because if triangles are required, we need construct new bounding planes for these new triangles, which violating our principle of no geometry construction (Fig. \ref{fig:iisect}d).

@@

The intersection vertices are not only introduced by intersection between an edge and a triangle, but also introduced by intersections of three triangles. In order to find all intersection vertices, it is not enough by only triangle-triangle intersection tests. To make tessellation easier and more robust, we have to refine these intersections and resolve the intersections between intersections.

Intersection refinement is performed in a local scope. For each intersecting face $t$ from mesh $M_t$, we collect all intersections on $t$ as set $\bm{\Gamma}(t)$
%$\bm{\Gamma}(t)=\{\bm{\mathcal{I}}_{t}^i, i=0,1,2,...\}\cup\{\bm{\mathcal{I}}_{t}^{\bm{e}_i},i=0,1,2\}$
and refine them together. For simpler processing, we also include the three edges of $t$ into $\bm{\Gamma}(t)$, since edges also participate in tessellation. In $\bm{\Gamma}(t)$, edges are represented as PBI. The neighborhood components of edges are set as $None$ because they do not have a neighborhood edge or face. Other PBI-rep components can be determined by the P-reps of edges. The refinement on $\bm{\Gamma}(t)$ is done using only plane-based geometry predicates by the following three steps:

\vspace{0.5em}
\noindent \textbf{Coincidence elimination}~~~~
We merge coincidence intersections which have the same end points. Intersections are undirected so intersections with inverse end points are also coincident. The PBI-rep of the new intersection is inherited from either of the old one's except the neighborhood component. The new neighborhood is the combination of the merged ones', meaning the intersection has two (or more) neighborhoods.

\begin{wrapfigure}{r}[0in]{0in}
\includegraphics[width=1.7 in]{resolve}
%\caption{Plane-based representation of triangle}
\end{wrapfigure}
\vspace{0.5em}
\noindent \textbf{Cross and overlap resolving}~~~~
After the process of two previous steps, there are no overlapping intersections in $\bm{\Gamma}(t)$. We check whether any pair of intersections cross or overlap (collinear) each other. The definition of \emph{cross} does not include the situation that two intersections share end points, since no splitting is required in this situation. When there is crossing or overlapping, at least one of the intersections have to be split. Because one intersection may may 'touch' more than one other intersections, the splitting is deferred until all splitting points are found. The subroutine linear order of points (\S\ref{sec:substrates}) is used to sort these splitting points along the splitting intersection. The PBI-reps of split segments inherit their father's except the third component which identifies the end points.

\vspace{0.5em}
\noindent \textbf{Coincidence elimination (again)}~~~~
Unfortunately, the overlap resolving may bring new coincident intersections. Therefore the first step is run again in the final step.

@@

So far, intersections are only independent data that store the coordinates and neighbouring information. To perform tessellation and extract subfaces, we need to organize them to reveal the topology. We use tess-graph for this purpose. A tess-graph is the graph description of the tessellated face topology. For each face to be tessellated, we construct a tess-graph according to the refined intersections. Nodes of tess-graph represent end points of intersections and undirectional connections between nodes represent intersections. The construction of tess-graph is straightforward and the reader will not have problem filling the details.

After we have tess-graph, we tessellate face and extract subfaces according to it. This is done by extract valid loops on tess-graph. A valid loop that corresponds to a subface must satisfy two criterions: 1) the direction of loops should be coherent with the face normal, and 2) consecutive connections on the loop should be adjacent by circular order. Each valid loop corresponds to a intersection-free face. After we find out all valid loops on tess-graph, we finish tessellating the corresponding face. To facilitate the later process, we also store neighborhood information into the edges of the new faces. Because the edges of new faces are generated from intersections, the neighborhood information of the edges is inherent from the corresponding intersections.

\begin{figure}[t]
\centering
\includegraphics[width=3in]{circularadj}
\caption{\emph{Left}: we find all valid loops on tess-graph to tesselate a triangle. The direction of loops have to be coherent with the triangle normal (assume the normal points to the outside of paper). \emph{Right}: each circular adjacent edge pair is an angle of the tessellated polygon.}
\label{fig:cadj}
\end{figure}

When all faces are tessellated, meshes are intersection-free. We further attach the PBI-reps of intersections to the edges generated from tess-graph connections which corresponds to an intersection. Finally, we get the intersection-free hybrid meshes, with some edges contains neighborhood information which indicates there are intersections between meshes.

@@

\label{sec:classification}
We traverse each face in the intersection-free meshes, and classify which one belongs to the final mesh. Classification is done by evaluating the indicator vector of each face. The basic idea of our classification method is to utilize the space coherence of face indicator vectors to reuse the classification results. Because we store the PBI-reps which contains neighborhood information in the edges which intersect other meshes, the search space of point-in-polyhedron test is limited to only a few faces, which makes the indicator computation very fast. Also, by using P-reps, our classification method can deal with any degenerate cases.

The space coherence of indicator vectors means neighboring faces could share the same indicator vector, or most components of indicator vectors. Therefore, we start from a seed face $\bm{s}_0$ with its indicator vector $\bm{\Lambda}(\bm{s}_0)$ known, and propagate to other faces in a flood-filling manner. The trace of indicator propagation is:
\begin{equation}
\label{eq:trace}
\bm{s}_0\to \bm{s}_1\to \bm{s}_2\to \bm{s}_3\to ...
\end{equation}

Our hybrid meshes facilitate this process by accelerating the indicator propagation. Given adjacent faces $\bm{s}_1$ and $\bm{s}_2$, it is straightforward to know whether $\bm{\Lambda}(\bm{s}_1)$ and $\bm{\Lambda}(\bm{s}_2)$ are different. The two indicator vectors are the same unless the shared edge $\bm{e}_{12}$ lies on the boundary of some other meshes, which means there is neighborhood information, say, $\bm{\mathcal{N}(\bm{e}_{12})} = \{(P_k, \mathcal{N}_k)\}$, stored on $\bm{e}_{12}$. The neighborhood information indicates which components of indicator vector differ between $\bm{s}_1$ and $\bm{s}_2$.
 \iffalse
 Namely, for a certain indicator $\lambda_k$ of $\bm{s}_1$ and $\bm{s}_2$, it always stands:
\begin{equation}
\label{eq:eq1}
{\lambda_k}(\bm{s}_1) = {\lambda_k}(\bm{s}_2),\ \ \  \mbox{if }\mathcal{N}_k(\bm{\mathcal{I}}(e_{12}))= 0.
\end{equation}
\fi
If there is a neighborhood $(P_k, \mathcal{N}_k) \in \bm{\mathcal{N}}(\bm{\bm{e}_{12}})$, the $k^{th}$ indicators $\lambda_k(\bm{s}_1)$ and $\lambda_k(\bm{s}_2)$ can be computed efficiently and exactly according to $\mathcal{N}_k$ (\S\ref{sec:propagation}). We outline our classification method in Algorithm \ref{code:floodfill}.

@@

Meshes consist of vertices, edges and faces. Classification is required to perform on the face level. However, polyhedron in-out test is performed based on points. Gaps exist between the them. Conventionally, face barycenter is used to compute the indicator of the whole face. This is because in intersection-free meshes, face is classified as a whole, and every face inner point will have the same indicator as the face's. However, the computation of inner point coordinates requires geometry construction, which introduces errors without exact arithmetic. Therefore, we use the vertices instead, whose exact representations are known, for indicator computation.

@@

%For example, given a face $\bm{s}$ and one of its vertex $v^i(\bm{s})$, if $\lambda_k(v^i(\bm{s}))=on$, $\lambda_k(\bm{s})$ can be any of $on$, $in$ or $out$ (see {\color{red}{Fig. ?}})
However, there are two gaps between face indicator and vertex indicator. First, the indicator of vertex does not always equal to indicator of face. Second, face indicator has two conditions in $on$ case---$same$ and $oppo$, because the face has its orientation.

To overcome the gaps, we discuss our indicator trace (\ref{eq:trace}) in finer granularity. In the beginning, we start from a seed vertex $\bm{v}_0$ on seed face $\bm{s}_0$. We assume indicator vector $\bm{\Lambda}(\bm{v}_0)$ is known. Then we use $\bm{\Lambda}(\bm{v}_0)$ to compute $\bm{\Lambda}(\bm{s}_0)$. In addition, we add the edge layer into the trace. Then it changes to:
\begin{equation}
\bm{v}_0\to \bm{e}_0\to \bm{s}_0\to (\bm{e}_1)\to \bm{s}_1\to (\bm{e}_2)\to \bm{s}_2\to ...
\end{equation}
, where $\bm{v}, \bm{e}$ means vertex and edge respectively. The bracket here means propagating to edge is optional because if the edge does not contain neighborhood information, the indicators on the two side are the same and we do not need to propagate to the edge. We find that there are three kinds of basic operations in the trace: $\bm{v}_{\bm{e}}\to \bm{e}$, $\bm{e}_{\bm{s}}\to \bm{s}$ and $\bm{s}\to \bm{e}_{\bm{s}}$, where $\bm{v}_{\bm{e}}$ means the end point of $\bm{e}$, and $\bm{e}_{\bm{s}}$ means the edge of $\bm{s}$.

\begin{figure}[t]
\centering
\includegraphics[width=3in]{bsps}
\caption{a) Because we cannot use inner sample point includes geometry construction, we choose the vertices (green one) instead to compute the face indicator. b-d): Different situations of classification. The simple structure of neighborhood BSP guarantees the same classification result of the points on $\bm{p}_{\bm{s}, sp}^+$. In \emph{a}, the BSP contains two half-planes because it is constructed from two neighborhood faces. In \emph{d}, the classification result is $on$ so we need to determine the orientation by the normal of half-plane $q$ and the face $p$. In this condition, the indicator should be $oppo$ since $p$ and $q$ are opposite.}
\label{fig:bsps}
\end{figure}

Given the partial order $on \succ in$ and $on \succ out$, our key observation is that the following relation always stands for indicators:
\begin{equation}
\label{eq:porder}
\lambda_k(\bm{v}_{\bm{e}_{\bm{s}}}) \succeq \lambda_k(\bm{e}_{\bm{s}}) \succeq \lambda_k(\bm{s})
\end{equation}
, where $\lambda_k(x)$ means the indicator of $x$ against a certain primitive $M_k$. This relation can be inferred from continuous space assumption and the definition of intersection-free mesh. It tells us when we perform $\bm{v}_{\bm{e}}\to \bm{e}$ and $\bm{e}_{\bm{s}}\to \bm{s}$, we decide whether the $on$ indicator changes to $in$ or $out$ or remains $on$. When we perform $\bm{s}\to \bm{e}_{\bm{s}}$, we decide whether $in$ and $out$ indicators change to $on$. In addition, considering the orientation of face, we also need to figure out whether the $on$ indicators changes to $same$ or $oppo$ during $\bm{e}_{\bm{s}}\to \bm{s}$ operation.

\vspace{0.5em}
\noindent\textbf{$\bm{s\to \bm{e}_{\bm{s}}}$}~~~~ We need to find which indicators change to $on$. We known when $\lambda_k(\bm{e}_{\bm{s}})=on$, there will be neighborhood information related to mesh $M_k$ stored on $\bm{e}_{\bm{s}}$. Therefore, in this operation, all we need is to iterate over neighborhoods of $\bm{e}_{\bm{s}}$ and change corresponding indicators to $on$.

>>>>

\IEEEtitleabstractindextext{
\begin{abstract}
  Constructive solid geometry (CSG) is widely used in computer aided design and manufacturing. However, boolean operations, which compute the boundary representations of CSG, have suffered from robustness and efficiency problems for more than three decades. We propose a fast and exact boolean operation method which is unconditionally robust for regular set solids. Previously reported boolean methods using plane representations (P-reps) of meshes are robust, but inefficient. Conversely, methods based on boundary representations (B-reps) of meshes are fast but not robust, unless exact arithmetic is applied. Therefore, we choose to combine P-reps with B-reps, which allowed us to use the advantages of both. Our method uses B-reps globally, for efficiency, then P-rep geometry is applied locally to ensure robustness. Comparison experiments with state-of-the-art techniques show that our method is unconditionally exact and robust, and similarly efficient to non-robust methods.
\end{abstract}

@@

ONSTRUCTIVE solid geometry (CSG) has long been a popular modeling tool for computer-aided design and computer-aided manufacturing (CAD/CAM). Complex models are constructed by combining primitives using a series of regularized boolean operations \cite{requicha1977mathematical,tilove1980closure}: union, intersection, and difference. CSG can be converted into the widely-used boundary representation (B-rep) through boolean operations.

@@

Another thing needs to be noticed is that there will be repetitive detection of edge intersections. For example, in Fig. \ref{fig:twin}a, the same intersection on $t_1$ will be detected twice because $t_1$ intersect both $t_2$ and $t_2^*$ in exactly the same intersection. We solve this duplication together with other interactions among different intersections in \S\ref{sec:tessellation}. We also call such $t_2$ and $t_2^*$ as \emph{companion triangles} in an edge intersection because they share the same edge intersection. This concept will be referred again in the discussion of coplanar cases.

@@

\begin{figure}[t]
\centering
\includegraphics[width=3.5in]{boolean-03}
\caption{a) Coplanar cases, $t_1$ and $t_2$ intersect in 2D, dividing each other into two areas---a convex overlapping area and exclusive parts. b) The possible configurations of the companion faces. The blue triangles are from the same mesh. For $\mathcal{I}_a$, the companion triangle $t_a$ is coplanar with the $t_1$. Therefore, the overlap areas ($t_1 \cap t_a$ and $t_1 \cap t_2$) do not need to be split by the yellow edge during tessellation, thus $\mathcal{I}_a$ is invalid. For $\mathcal{I}_b$ and $\mathcal{I}_c$, the companion faces are not coplanar, so these intersection can be recorded during intersection test between $t_1$ and $t_a$ or $t_1$ and $t_b$. In any situation, there is no need to record the intersection between $t_1$ and $t_2$.}
\label{fig:coplanar}
\end{figure}

Consider two triangle faces $t_1\in M_x$ and $t_2 \in M_y$ intersect within a common plane. Both $t_1$ and $t_2$ will divide each other into two areas---a convex overlapping area and exclusive parts (Fig. \ref{fig:coplanar}a). Apparantly, if we tessellate both $t_1$ and $t_2$ according to the boundary of overlapping area, we can guarantee that the tessellated meshes are intersection-free. Many previous methods \cite{feito2013fast,zhou2016mesh} use this method to guarantee topology correctness. However, in our method, we do not test whether $t_1$ and $t_2$ really intersect in 2D once we find they are coplanar. We treat coplanar situations as they do not intersect at all. In this way, we simplify our method, making it more robust and fast, while doing no harm to the topology correctness.

@@

However, if the companion triangles are also coplanar, neither of the two triangles will record the intersection. Fortunately, in this case, the intersection is not valid. The word `valid` means the intersections will appear as an edge in the final mesh, thus is necessary to put into tessellation. Supposing the intersection $\mathcal{I}$  is on $t_1$, this definition equals to:
\begin{equation}
% = f(\bm{\Lambda}(\bm{s}_i))
\lambda_f(\bm{v}_i) \neq constant, \bm{v}_i \in \bm{U}(\mathcal{I}) \cap M_x
\end{equation}
, where $f$ is the CSG function, and $\bm{U}(\mathcal{I})$ is the the neighborhood of $\mathcal{I}$ in $\mathbb{R}^3$. This inequation indicates the space near $\mathcal{I}$ on mesh $M_x$ have different classification and must be split according to $\mathcal{I}$. However, in most time, the validation of intersection is not known until classification stage, because we have to know the indicator of each point in $ \bm{U}(\mathcal{I}) \cap M_x$. Fortunately, if companion face of $t_2$ is coplanar with $t_2$, both sides of $\mathcal{I}$ on $t_1$ have the same indicator $\lambda_y=on$ and thus the same $\lambda_f$. It means $\mathcal{I}$ is invalid and we do not need to record it.

@@

\label{sec:tessellation}
After intersection computation, tessellation is performed on each intersecting triangle face to generate the intersection-free meshes. We call this step as deferred tessellation because the tessellation happens until all valid intersections are computed, instead of clipping triangle incrementally by each intersection.

Ogayar-Anguita et. al. \cite{ogayar2015deferred} use Constrained Delaunay Triangulation (CDT) to perform deferred tessellation. They treat triangle faces as convex triangulation zones and intersections as constraints. Our first try is to implement a robust plane-based CDT. However, CDT algorithms \cite{chew1989constrained,preparata2012computational} require coordinates projection or intersection detection between arbitrary connections of vertices within triangulation zone, which are difficult to be implemented under P-reps. From another perspective, CDT contains geometry constructions---new edges are generated to guarantee each subface is a triangle (see Fig. \ref{fig:iisect}d), which is not preferred in our method. Moreover, when there are more than two primitives, simply applying CDT omits the complexity of intersections between meshes---the intersections may intersect each other, introducing new vertices and splitting original intersection line segments (see Fig. \ref{fig:iisect}a). This breaks the assumption of most CDT algorithms and may lead to incorrect output topology.

\begin{figure}[t]
\centering
\includegraphics[width=3.5in]{boolean-04}
\caption{a) Different colors indicate the intersections are from different meshes. The yellow intersection and the red intersection intersect at a point. Also, the yellow intersection overlap with the green intersection. b) After refinement, we introduce a new vertex $\bm{v}_a$, and merge overlapping intersections into one edge $\bm{e}_b$. c) Our tessellation does not guarantee to obtain triangles. d) If triangulation is performed, new edges (red lines) are introduced and we cannot give the precise plane-based representation for these new edges.}
\label{fig:iisect}
\end{figure}

To solve these problems, we first perform intersection refinement to eliminate any crossing or overlapping between intersections. After that, we perform conservative tessellation based on \emph{tess-graph}, ensuring unconditional topology correctness. An tess-graph is a graph-like description of intersections on a certain face. The word `conservative` means we do not add any new edge during tessellation, which do not guarantee the subfaces are triangles.

@@

\label{sec:refine}
The intersection vertices are not only introduced by intersection between an edge and a triangle, but also introduced by intersections of three triangles. In order to find all intersection vertices, it is not enough by only triangle-triangle intersection tests. Thus, we have to refine these triangle-triangle intersections before the real tessellation.

Intersection refinement is performed in a local scope. For each intersecting face $t$, we collect all intersections on $t$ as a set $\bm{\Gamma}(t)$ and refine them together. We also include the three edges of $t$ into $\bm{\Gamma}(t)$, since edges also participate in tessellation. In $\bm{\Gamma}(t)$, edges are represented as PBI: the neighborhood $\mathcal{N}$ of edges are set as $N/A$ because they do not have a neighborhood. Other PBI-rep components can be determined by the P-reps of edges. The refinement on $\bm{\Gamma}(t)$ is done using only plane-based geometry predicates by the following three steps:

\vspace{0.5em}
\noindent \textbf{Coincidence elimination}~~~~
We merge coincidence intersections which have the same end points. Intersections are undirected so intersections with inverse end points are also coincident. The PBI-rep of the merging intersection is inherited from either of the old one's except the neighborhood component. The new neighborhood is the union of the merged ones', meaning the intersection has two (or more) neighborhoods.

\begin{wrapfigure}{r}[0in]{0in}
\includegraphics[width=1.7 in]{resolve}
%\caption{Plane-based representation of triangle}
\end{wrapfigure}
\vspace{0.5em}
\noindent\textbf{Cross and overlap resolving}~~~~
After the process of two previous steps, there are no overlapping intersections in $\bm{\Gamma}(t)$. We check whether any pair of intersections cross or overlap (collinear) each other. New vertices might be introduced in this step. The definition of \emph{cross} does not include the situation that two intersections share one end point, since no splitting is required in this situation. When there is crossing or overlapping, at least one of the intersections have to be split. Because one intersection may may 'touch' more than one other intersections, the splitting is deferred until all splitting points are found. The subroutine linear order of points (\S\ref{sec:substrates}) is used to sort these splitting points along the splitting intersection. The PBI-reps of split segments inherit their father's except the $\bm{P}_0, \bm{P}_1$, which identifies the P-reps of end points*.

\vspace{0.5em}
\noindent \textbf{Coincidence elimination (again)}~~~~
Unfortunately, the overlap resolving may bring new coincident intersections in overlap cases. Therefore the first step is run again in the final step.

@@

So far, intersections are only independent data that store the coordinates and neighbourhood information. To perform tessellation and extract subfaces, we need to organize them to reveal the topology. We use tess-graph for this purpose. A tess-graph is the graph description of the tessellated face topology. For each face to be tessellated, we construct a tess-graph according to the refined intersections. Nodes of tess-graph represent end points of intersections and undirectional connections between nodes represent intersections. The construction of tess-graph is straightforward and the reader will not have problem filling the details.

After we have tess-graph, we tessellate face and extract subfaces according to it. This is done by extract valid loops on tess-graph. A valid loop must satisfy two criterions: 1) the direction of loops should be coherent with the face normal, and 2) consecutive connections on the loop should be adjacent by circular order. Each valid loop corresponds to a intersection-free face. After we find out all valid loops on tess-graph, we finish tessellating the corresponding face (see Fig. \ref{fig:cadj}). To facilitate the later process, we also store neighborhood information into the edges of the new faces. The neighborhood information is inherent from the corresponding intersections. When all faces are tessellated, meshes are intersection-free.

\begin{figure}[t]
\centering
\includegraphics[width=3in]{boolean-05}
\caption{a) We find all valid loops on tess-graph to tesselate a triangle. The direction of loops have to be coherent with the triangle normal (assume the normal points to the outside of paper). b) Each circular adjacent edge pair is an angle of the tessellated polygon.}
\label{fig:cadj}
\end{figure}

@@

\label{sec:classification}
We traverse each face in the intersection-free meshes, and classify which one belongs to the final mesh. Classification is done by evaluating the indicator vector $\bm{\Lambda}$ of each face. The basic idea of our classification method is to utilize the space coherence of face indicator vectors to share the classification results among neighboring faces. Also, because we store the neighborhood information in the edges which intersect other meshes, the search space of point-in-polyhedron test is limited to only a few faces. It makes the indicator computation very fast. Finally, by using P-reps, our classification method is exact can deal with any degenerate cases.

The space coherence of indicator vectors means neighboring faces may have the same indicator vector, or most components of indicator vectors. Therefore, we start from a seed face $\bm{s}_0$ with its indicator vector $\bm{\Lambda}(\bm{s}_0)$ known, and propagate to other faces in a flood-filling manner. The trace of indicator propagation is:
\begin{equation}
\label{eq:trace}
\bm{s}_0\to \bm{s}_1\to \bm{s}_2\to \bm{s}_3\to ...
\end{equation}

Our intersection-free meshes facilitate this process by accelerating the indicator propagation. Given adjacent faces $\bm{s}_1$ and $\bm{s}_2$, it is straightforward to know whether $\bm{\Lambda}(\bm{s}_1)$ and $\bm{\Lambda}(\bm{s}_2)$ are different. The two indicator vectors are the same unless the shared edge $\bm{e}_{12}$ lies on the boundary of some other meshes, which means there is neighborhood information, say, $\bm{\mathcal{N}}(\bm{e}_{12}) = \{\mathcal{N}_k\}$, stored on $\bm{e}_{12}$. Noted that here we use $\bm{\mathcal{N}}$ to indicate that there may be more than one neighborhood for the edge (see \S\ref{sec:refine}). The neighborhood information indicates which components of indicator vector differ between $\bm{s}_1$ and $\bm{s}_2$.
 \iffalse
 Namely, for a certain indicator $\lambda_k$ of $\bm{s}_1$ and $\bm{s}_2$, it always stands:
\begin{equation}
\label{eq:eq1}
{\lambda_k}(\bm{s}_1) = {\lambda_k}(\bm{s}_2),\ \ \  \mbox{if }\mathcal{N}_k(\bm{\mathcal{I}}(e_{12}))= 0.
\end{equation}
\fi
If there is a neighborhood from mesh $M_x$, the $x^{th}$ indicators $\lambda_x(\bm{s}_1)$ and $\lambda_x(\bm{s}_2)$ differ and can be computed efficiently and exactly according to the neighborhood. We outline our classification method in Algorithm \ref{code:floodfill}.

@@

Meshes consist of vertices, edges and faces. Classification is required to perform on the face level. However, polyhedron-in-out test is performed based on points. Gaps exist between the them. Conventionally, face barycenter is used to compute the indicator of the whole face. This is because in intersection-free meshes, face is classified as a whole, and every face inner point will have the same indicator as the face's. However, the computation of inner point coordinates requires geometry construction, which introduces errors without exact arithmetic. Therefore, we use the face (or subface) vertices instead, whose exact representations are known, for indicator computation.

@@

However, there are two gaps between face indicator and face vertex indicator. First, the indicator of vertex does not always equal to indicator of face. Second, face indicator has two extra classes in $on$ case---$same$ and $oppo$, because the face has its orientation.

To overcome the gaps, we rewrite our indicator trace (\ref{eq:trace}) in finer granularity. In the beginning, we start from a seed vertex $\bm{v}_0$ on seed face $\bm{s}_0$. We assume indicator vector $\bm{\Lambda}(\bm{v}_0)$ is known. Then we use $\bm{\Lambda}(\bm{v}_0)$ to compute $\bm{\Lambda}(\bm{s}_0)$. In addition, we add an optional edge layer into the trace. Then it changes to:
\begin{equation}
\bm{v}_0\to \bm{e}_0\to \bm{s}_0\to (\bm{e}_1)\to \bm{s}_1\to (\bm{e}_2)\to \bm{s}_2\to ...
\end{equation}
, where $\bm{v}, \bm{e}$ means vertex and edge respectively. The bracket here means propagating to edge is optional because if the edge does not contain neighborhood information, the indicators on the two side are the same and we do not need to propagate to the edge. We find that there are three kinds of basic operations in the trace: $\bm{v}_{\bm{e}}\to \bm{e}$, $\bm{e}_{\bm{s}}\to \bm{s}$ and $\bm{s}\to \bm{e}_{\bm{s}}$, where $\bm{v}_{\bm{e}}$ means the end point of $\bm{e}$, and $\bm{e}_{\bm{s}}$ means the edge of $\bm{s}$.

\begin{figure}[t]
\centering
\includegraphics[width=3in]{bsps}
\caption{a) Because inner sample point includes geometry construction, we choose the face vertex (orange one) instead to compute the face indicator. b-d): Different situations of classification. The simple structure of neighborhood BSP guarantees the same classification result of the points on $\bm{p}_{\bm{s}, sp}^+$. In d), the classification result is $on$ so we need to determine the orientation by the normal of BSP splitting plane $q$ and the face normal $p$. In this condition, the indicator should be $oppo$ since $p$ and $q$ are opposite.}
\label{fig:bsps}
\end{figure}

Given the partial order $on \succ in$ and $on \succ out$, our key observation is that the following relation always stands for indicators within intersection-free meshes:
\begin{equation}
\label{eq:porder}
\lambda_k(\bm{v}_{\bm{e}_{\bm{s}}}) \succeq \lambda_k(\bm{e}_{\bm{s}}) \succeq \lambda_k(\bm{s})
\end{equation}
, where $\lambda_k(x)$ means the indicator of $x$ against a certain primitive $M_k$. This relation can be inferred from continuous space assumption and the definition of intersection-free mesh*. It tells us when we perform $\bm{v}_{\bm{e}}\to \bm{e}$ and $\bm{e}_{\bm{s}}\to \bm{s}$, from low-dimension to high-dimension, we decide whether the $on$ indicator changes to $in$ or $out$ or remains $on$. When we perform $\bm{s}\to \bm{e}_{\bm{s}}$, we decide whether $in$ and $out$ indicators change to $on$. In addition, considering the orientation of face, we also need to figure out whether the $on$ indicators changes to $same$ or $oppo$ during $\bm{e}_{\bm{s}}\to \bm{s}$ operation.

\vspace{0.5em}
\noindent\textbf{$\bm{s\to \bm{e}_{\bm{s}}}$}~~~~ We need to find which indicators change to $on$. We known when $\lambda_k(\bm{e}_{\bm{s}})=on$, there will be neighborhood information related to mesh $M_k$ stored on $\bm{e}_{\bm{s}}$. Therefore, in this operation, all we need is to traverse over all neighborhoods of $\bm{e}_{\bm{s}}$ and change corresponding indicators to $on$.

\vspace{0.5em}
\noindent\textbf{$\bm{\bm{e}_{\bm{s}}\to \bm{s}}$}~~~~From eq. \ref{eq:porder}, we know if $\lambda_k(\bm{e}_{\bm{s}}) \neq on$, then $\lambda_k(\bm{e}_{\bm{s}})=\lambda_k(\bm{s})$. On the other hand, when $\lambda_k(\bm{e}_{\bm{s}}) = on$, we figure out $\lambda_k(\bm{s})$ using neighborhood information related to $M_k$. The neighborhood related to $M_k$ can be either a face or a set of faces from mesh $M_k$.  We can build a trivial BSP \cite{thibault1987set} with these faces. The BSP can be used to compute $\lambda_k(\bm{s})$ if we can sample a point from $\bm{s} \cap \bm{U}(\bm{e}_{\bm{s}})$, where $\bm{U}(\bm{e}_{\bm{s}})$ is the neighborhood space of $\bm{e}_{\bm{s}}$. Unfortunately, we cannot guarantee to find such a point with precise coordinates.

However, we notice that our BSP is constructed from a single face or a set of faces sharing the same edge. Such BSP is so simple that indicators $\lambda_k$ of all points on the half plane $\bm{p}_{\bm{s}, sp}^+$ are the same. Here $\bm{p}_{\bm{s}, sp}^+$ is half of the supporting plane of $\bm{s}$ that on the positive side of bounding plane $\bm{p}_{\bm{s}, b}^{\bm{e}_{\bm{s}}}$ (see Fig. \ref{fig:bsps}). Since $\bm{p}_{\bm{s}, sp}^+ \cap \bm{s} \neq \varnothing$, we can compute $\lambda_k(\bm{s})$ by figuring out $\lambda_k(\bm{v}_x, \bm{v}_x \in \bm{p}_{\bm{s}, sp}^+)$. For polygons, there is at least one corner point on $\bm{p}_{\bm{s}, sp}^+$, whose exact coordinates are known. Then that corner point is used as the $\bm{v}_x$. Also, we assign each BSP splitting plane a normal by the normal of triangle it contains, by which we can decide the orientation ($same$ or $oppo$) when $\lambda_k(\bm{s})=on$.
